Mean Estimation from One-bit Measurements

### 2017-10-05 Slides/OneBit_Allerton_slides.tex
new version of presentation

\usepackage{pgfplots}
\usepackage{graphicx}
\usepackage{pstool}

@@

\newcommand{\mmse}{\mathsf{mmse}}
\newcommand{\thetac}{{\color{red} \theta}}
\newcommand{\enc}{\mathsf{enc}}
\newcommand{\qnt}{\mathsf{qnt}}
\newcommand{\sgn}{\mathsf{sign}}
\newcommand{\unif}{\mathsf{unif}}

@@

\subsection{Motivation}
\begin{frame}
\frametitle{Motivation}
\framesubtitle{}
Point estimation under communication constraints:\\
\begin{center}
\begin{tikzpicture}[>= latex', node distance = 2cm]
\node at (0,0) (p_th) {$p_{\thetac}(x)$};
\node[right of = p_th] (sample) {$X_1,\ldots,X_n$};
\node[box, right = .75cm of sample, align = left](pm) {limited bits \\ per sample
};
\node[right = .75 cm of pm](th_hat) {$\widehat{\thetac}$};
%\node[below = .1cm of  Xhat]{\textcolor{blue}{reconstruction}};
\draw[->, thick] (p_th) -- (sample);
\draw[->, thick] (sample) -- (pm);
\draw[->, thick] (pm) -- (th_hat);
\end{tikzpicture}
\end{center}
Estimation error is due to:
\begin{itemize}
\item[(i)] limited data 
\item[(ii)] limited bits
\end{itemize}
\bigskip

Relevant scenarios:
\begin{itemize}
\item big data
\item low-power sensors 
\item distributed computing / optimization (bottleneck is due to communication between processing units)

@@

\node[below of = source2, node distance = 2.5cm] (source3) {$X_n$};
\node[int1, right of = source3, node distance = 2cm, scale = 0.75] (enc3) {Enc n};  
\draw[->,line width = 2pt] (source3) -- (enc3);

%\node[above of = source, node distance = 1cm] (dist) {${\mathcal N} \left(\theta, \sigma^2 \right)$};

\node[Est, right of = enc2, node distance = 5cm] (est) {Est};
\draw[->,line width = 2pt] (enc1) -| node[above, xshift = -2.5cm] (mes1) {$M_1 \in \left\{-1,1\right\}$} (est);   
\draw[->,line width = 2pt] (enc2) -- node[above, xshift = 0cm] (mes2) {$M_2 \in \left\{-1,1\right\}$} (est);   
\draw[->,line width = 2pt] (enc3) -| node[above, xshift = -2.5cm]  {$M_n \in \left\{-1,1\right\}$} (est);   
\node[right of = est, node distance = 2.5cm] (dest) {$\widehat{\theta}(M^n)$};
\draw[->, line width=2pt] (est) -- (dest);

@@

\visible<2>{
\draw[<->,line width = 2pt, color = red] (enc1) -- (enc2);   
\draw[<->,line width = 2pt, color = red] (enc2) -- (enc3);
\draw[dashed,line width = 2pt, color = white] (enc2)+(0,-0.5) -- +(0,-1.5);}

\visible<3->{
\draw[->,line width = 2pt, color = blue] (enc1) -- (enc2);   
\draw[->,line width = 2pt, color = blue] (enc2) -- (enc3);
\draw[dashed,line width = 2pt, color = white] (enc2)+(0,-0.7) -- +(0,-1.7);}

@@

\begin{itemize}
\item Distributed: $M_i = f_i(X_i)$
\visible<2->{
\item {\color{red} Centralized}: $M^n = (M_1,\ldots,M_n) = f(X_1,\ldots,X_n)$ }
\visible<3->{
\item {\color{blue} Adaptive / Sequential}: $M_i = f_i(X_i,M^{i-1})$ }
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Related Work}
\begin{itemize} 
\item Estimation via compressed information [Han '87], [Zhang \& Berger '88]
\item Distributed hypothesis testing under quantization [Tsitsiklis '88]
\item Estimation from multiple machines subject to a bit constraint [Zhang, Duchi, Jordan, Wainwright '13]
\item Remote multiterminal source coding (CEO) [Berger, Zhang, Wiswanathan '96], [Oohama '97]
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Consistency}
Q: in what setting consistent estimation is possible? \\
\bigskip
\pause
A: all !
\[
M_i = \mathbf 1(X_i>0),\quad i=1,\ldots,n
\]
(as in the distributed setting)
\[
\frac{1}{n}\sum_{i=1}^n M_i \rightarrow \Prob\left(X<0 \right) = \Phi\left(X/\sigma<\thetac\right)
\]
\end{frame}

@@

\begin{frame}
\frametitle{Efficiency}
\textbf{Definition:} \emph{asymptotic relative efficiency (ARE)} of an estimator:
\[
\mathrm{ARE}(\widehat{\theta}) \triangleq \lim_{n\rightarrow \infty} \frac{\mathbb E \left[ \left(\widehat{\theta}-\theta \right)^2 \right]}{\sigma^2 / n}
\]
($\sigma^2/n$ is the minimax risk without communication constraints)
%MSE of sample mean $\bar{\theta}$: 
%\[
%\mathbb E \left[ \left(\bar{\theta}-\theta \right)^2 \right] = \frac{\sigma^2}{n}
%\]
%(minimax estimator without communication constraint)

\bigskip
\pause
Q: in what scenarios the ARE is finite? \\
\bigskip
\pause
This talk: all three scenarios

@@

\begin{prop} If the parameter space $\Theta$ is bounded, then the ARE under centralized encoding is $1$
\end{prop}
\textbf{Proof:}
\[
\mathbb E \left( \theta- \widehat{\theta} \right)^2 = \overbrace{\mathbb E \left( \theta- \bar{\theta} \right)^2}^{\sigma^2/n} + \mathbb E \left( \bar{\theta}- \widehat{\theta} \right)^2 
\]
\begin{itemize}
\item Encoder is required to describe $\bar{\theta}$ using $n$ bits
\begin{itemize}
\item divide parameter space $\Theta$ into $2^n$ regions of equal size
\item send region index where $\bar{\theta}$ falls
\end{itemize}
\item MSE in estimating $\bar{\theta}$ decreases exponentially in $n$ 
\end{itemize}
\bigskip
Note: globally optimal strategy for a finite $n$ is hard to derive since mean of $\bar{\theta}$ is unknown 
\end{frame}

@@

\begin{center}
\begin{tikzpicture}[node distance=2cm,auto,>=latex]
  \node at (0,0) (source) {$\thetac$} ;
  \node[ssum, right of = source, scale = 0.75] (sum1) {+};
  \node[above of = sum1, node distance = 1.2cm] (z1) {$Z_1$};
  \node[int1, right of = sum1, node distance = 2cm, scale = 0.75] (enc1) {Enc 1};  
\draw[->,line width = 2pt] (source) -- (sum1); 
\draw[->,line width = 2pt] (sum1) -- node[above] {$X_1$} (enc1); 
\draw[->,line width = 2pt] (z1) -- (sum1);

@@

 \node[coordinate,below of = source, node distance = 2.5cm] (source2) {};
 \node[ssum, right of = source2, scale = 0.75] (sum2) {+};
  \node[above of = sum2, node distance = 1.2cm] (z2) {$Z_n$};
\node[int1, right of = sum2, node distance = 2cm, scale = 0.75] (enc2) {Enc n};

@@

\draw[->,line width = 2pt] (sum2) -- node[above] {$X_n$} (enc2); 
\draw[->,line width = 2pt] (z2) -- (sum2);

@@

%\node[int1, right of = enc2, node distance = 6cm, scale = 0.75] (est) {Est};
\draw[->,line width = 2pt] (enc1) -- node[above, xshift = 0cm] (mes1) {$R_1$} (est);   
\draw[->,line width = 2pt] (enc2) -- node[above, xshift = 0cm] (mes2) {$R_n$} (est);   
\node[right of = est, node distance = 2cm] (dest) {$\widehat{\theta}$};
\draw[->, line width=2pt] (est) -- (dest);
\end{tikzpicture}
\end{center}

Assume:
\begin{itemize}
\item $\theta \sim \mathcal N(0,\sigma_\theta^2)$, ~~~ 
\item $Z_1,\ldots,Z_n \overset{\mathrm{i.i.d}}{\sim}\mathcal N(0,\sigma^2)$
\end{itemize}
Encode $k$ instances:
\begin{itemize}
\item $R_1 = \ldots = R_n =1 $
\item $D_{CEO} = \frac{1}{k} \sum_{j=1}^k \mathbb E \left(\theta_j-\widehat{\theta}_j \right)^2$
\end{itemize}
\end{frame}

\begin{frame}

From [K., Rini, Goldsmith '17]: 
\[
D_{CEO} \leq \frac{4}{3} \frac{\sigma^2}{n} + o(1)
\]
%\begin{itemize}
%\item[] CEO distortion satisfies [Chen, Zhang, Berger, Wicker '04]:
%\[ D^\star = \frac{4}{3} %\frac{\sigma^2}{n} + o(n^{-1}) \]
%\end{itemize}
ARE of $4/3$ can be attained in a fully distributed encoding (with encoding over blocks of multiple problem instances)
\bigskip
\begin{alertblock}{Conclusion}
Distributed encoding is almost not a limiting factor 
(although inability to exploit concentration of measure in high dimension -- might!)
\end{alertblock}
\end{frame}

@@

@@

@@

\begin{theorem}[achievability]
Assume that $\Theta$ is a bounded interval. There exists an estimator with ARE equals to $\pi/2$
\end{theorem}

\begin{theorem}[one-step optimal strategy]
The next step one-bit message that minimizes the MSE is of the form $M = \sgn (X-\tau)$ where $\tau$ satisfies the fixed-point equation 
\[
\tau = \frac{1}{2} \left( \frac{ \int_{-\infty}^\tau \theta \pi(d\theta) }{\int_{-\infty}^\tau \pi(d\theta)}  + \frac{ \int_{\tau}^\infty \theta \pi(d\theta)}{\int_{\tau}^\infty \pi(d\theta)} \right)
\]
\end{theorem}
\end{frame}

\begin{frame}
\frametitle{Proof}
\framesubtitle{converse (ARE $\geq \pi/2$)}
Assume a prior $\pi(d\theta)$ on $\Theta$ with location Fisher information $I_\pi$. \\

The van-Trees inequality (e.g. [Tsybakov '08]) implies
\[
\mathbb E \left(\theta - \widehat{\theta} \right)^2 \geq \frac{1}{\mathbb E I_\theta(M^n) + I_\pi} \geq \frac{1}{\sum_{i=1}^n I_\theta(M_i|M^{i-1}) + I_\pi} 
\]
\begin{lem}
\[
I_\theta(M_i|M^{i-1}) \leq 2/(\pi\sigma^2)
\]
\end{lem}
(proof by induction over a finite set of intervals approximating $M_i^{-1}(1)$ given $M^{i-1}$)
\end{frame}

\begin{frame}
\frametitle{Proof}
\framesubtitle{achievability (existence of an estimator with ARE $= \pi/2$)}
[Polyak \& Juditsky '92]: \\
\[
\begin{cases}
\theta_i = \theta_{i-1} + \gamma_i \varphi(X_i - \theta_i)& i=1,\ldots,n\\
\widehat{\theta} = \frac{1}{n} \sum_{i=1}^n \theta_i &
\end{cases}
\]
where:
\begin{itemize}
\item[(i)] $\gamma_n \rightarrow 0^+$ ``not too slow'' 
\item[(ii)] $\psi(x) = \mathbb E \varphi(x+Z)$ 
\item [(iii)] $\chi(x) = \mathbb E \varphi^2(x+Z)$ 
\item[(iv)] some regularity  conditions on $\varphi$, $\chi$, $\psi$
\end{itemize}
Then 
\[
\sqrt{n}(\theta- \widehat{\theta}) \rightarrow \mathcal N\left(0,V\right)
\]
where $V = \chi(0)/\psi'^2(0)$.  \\
\medskip
\pause
Proof of theorem: take $\varphi(x) = \sgn(x)$
\end{frame}

\begin{frame}
\frametitle{One-step optimality}
%Upon observing $M^{i-1}$, what is the optimal detection region $M_i^{-1}(1)$ that minimizes the MSE ? \\
\begin{theorem}
Let $\pi(\theta)$ be an absolutely continuous log-concave distribution. For $X\sim \mathcal N(\theta,\sigma^2)$ let
\[
M = \sgn(X-\tau)
\]
where $\tau$ is the unique solution to
\[
\tau = \frac{1}{2} \left( \frac{ \int_{-\infty}^\tau \theta \pi(d\theta) }{\int_{-\infty}^\tau \pi(d\theta)}  + \frac{ \int_{\tau}^\infty \theta \pi(d\theta)}{\int_{\tau}^\infty \pi(d\theta)} \right)
\]
Then for any $M'(X) \in \{-1,1\}$ and $\widehat{\theta}(M')$:
\[
\mathbb E \left(\theta- \widehat{\theta}(M') \right)^2 \geq \mathbb E \left(\theta - \mathbb E[\theta|M] \right)^2
\]
\end{theorem}
\vspace{-20pt}
\begin{alertblock}{Interpertation:}
The optimal one-bit message is a threshold detector. The threshold is the fixed-point that balances conditional center of masses given the message
\end{alertblock}

\begin{frame}
\frametitle{One-step Optimal Scheme}
Initialization: $P_0(t) = \pi(\theta)$ \\
Repeat for $n \geq 1$:
\begin{itemize}
\item[(i)] $P_n(t) = \Prob(\theta = t|M^n) = \alpha_n P_{n-1}(t) \Phi \left(M_n \frac{t-\tau_{n-1}}{\sigma} \right)$ 
\item[(ii)] $\widehat{\theta} = \mathbb E[\theta | M^n] = \int t P_n(t)dt $ 
\item[(iii)] Find $\tau_n$ from 
\[
\tau_n = \frac{1}{2} \left( \frac{ \int_{-\infty}^\tau t P_n(t)dt }{\int_{-\infty}^\tau P_n(t)dt}  + \frac{ \int_{\tau}^\infty t P_n(t)dt}{\int_{\tau}^\infty P_n(t)dt} \right)
\]
\item[(iv)] $M_{n+1} = \sgn(X_{n+1}-\tau_n)$
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Numerical Example}
\begin{center}
Normalized empirical risk versus number of samples $n$ \\
(500 Monte Carlo experiments)
\begin{tikzpicture}
\node at (0,0) {\includegraphics[scale=0.5]{one_bit_adpative}};
%\node[color = blue,  minimum width=1cm, align = left] at (1.5,1.5) {assymptotically optimal \eqref{eq:sgd_est}};
%\node[color = red, minimum width=1cm, align = left] at (1.5,0.55) {one-step optimal \eqref{eq:message_update}} ;
\node[rotate = 90, scale = 1] at (-5,0) {$n \mathbb E \left(\widehat{\theta}_n - \theta \right)^2$};
\node[scale = 1] at (0,-3) {$n$};
%\draw[dashed] (-3.5,1.18) node[left, scale = 0.7] {$\frac{\pi}{2}$}-- (-3,1.18);
\end{tikzpicture}
$\theta \sim \unif(-3,3)$
\end{center}
\end{frame}

@@

\begin{frame}
\frametitle{Summary}
\begin{itemize}
\item Asymptotic relative efficiency in adaptive setting is $\pi/2$ regardless of size of parameter space -- only $\sim 1.57$ more samples are required due to 1-bit constraints
\item One-step optimal one-bit message is a threshold detector
\item ARE in distributed setting is finite
\item 
ML estimator is local asymptotically optimal for threshold detection
\item ARE of ML is characterized by asymptotic density of threshold values
\item Minimax ARE of ML depends on size of parameter space
\end{itemize}

\bigskip
\begin{alertblock}{Open question}
Is there a distributed encoding scheme with ARE that is both finite and independent of size of parameter space ?
\end{alertblock}

>>>>

\usepackage{pgfplots}
\usepackage{graphicx}
\usepackage{pstool}
%\usepackage{beamerthemesplit}
\usepackage{stmaryrd}

@@

\newcommand{\mmse}{\mathsf{mmse}}
\newcommand{\thetac}{{\color{red} \theta}}
\newcommand{\enc}{\mathsf{enc}}
\newcommand{\qnt}{\mathsf{qnt}}
\newcommand{\sgn}{\mathsf{sign}}
\newcommand{\unif}{\mathsf{unif}}
\newcommand{\iid}{\mathrm{i.i.d.}}

@@

\subsection{Motivation}
\begin{frame}
\frametitle{Motivation}
\framesubtitle{}
Point estimation under communication constraints:\\
\begin{center}
\begin{tikzpicture}[>= latex', node distance = 2cm]
\node at (0,0) (p_th) {$p_{\thetac}(x)$};
\node[right of = p_th] (sample) {$X_1,\ldots,X_n$};
\node[box, right = .75cm of sample, align = left](pm) {limited bits \\ per sample
};
\node[right = .75 cm of pm](th_hat) {$\widehat{\thetac}$};
%\node[below = .1cm of  Xhat]{\textcolor{blue}{reconstruction}};
\draw[->, thick] (p_th) -- (sample);
\draw[->, thick] (sample) -- (pm);
\draw[->, thick] (pm) -- (th_hat);
\end{tikzpicture}
\end{center}
\pause
Estimation error is due to:
\begin{itemize}
\item[(i)] limited data 
\item[(ii)] limited bits
\end{itemize}
\bigskip
\pause
Relevant scenarios:
\begin{itemize}
\pause
\item big data
\pause
\item low-power sensors 
\pause
\item distributed computing / optimization % (bottleneck is due to communication between processing units)

@@

\node[below of = source2, node distance = 2cm] (source3) {$X_n$};
\node[int1, right of = source3, node distance = 2cm, scale = 0.75] (enc3) {Enc n};  
\draw[->,line width = 2pt] (source3) -- (enc3);

\node[above of = source, node distance = 0.75cm] (dist) {$X_i \sim {\mathcal N} \left(\theta, \sigma^2 \right)$};

\node[Est, right of = enc2, node distance = 4.5cm] (est) {Est};
\draw[->,line width = 2pt] (enc1) -| node[above, xshift = -2.5cm] (mes1) {$M_1 \in \left\{-1,1\right\}$} (est);   
\draw[->,line width = 2pt] (enc2) -- node[above, xshift = 0cm] (mes2) {$M_2 \in \left\{-1,1\right\}$} (est);   
\draw[->,line width = 2pt] (enc3) -| node[above, xshift = -2.5cm]  {$M_n \in \left\{-1,1\right\}$} (est);   
\node[right of = est, node distance = 2cm] (dest) {$\widehat{\theta}(M^n)$};
\draw[->, line width=2pt] (est) -- (dest);

@@

\visible<2>{
\draw[<->,line width = 2pt, color = blue] (enc1) -- (enc2);   
\draw[<->,line width = 2pt, color = blue] (enc2) -- (enc3);
\draw[dashed,line width = 2pt, color = white] (enc2)+(0,-0.5) -- +(0,-1.5);}

\visible<3->{
\draw[->,line width = 2pt, color = red] (enc1)+(1,0) -- (enc2);   
\draw[->,line width = 2pt, color = red] (enc2)+(1,0) -- (enc3);
\draw[line width = 2pt, color = red] (enc1)+(1,0) -- +(1,-1.5);

\draw[dashed,line width = 2pt, color = white] (enc2)+(0.66,-0.7) -- +(0.25,-1.5);}

@@

\begin{itemize}
\item Distributed: $M_i = f_i(X_i)$
\visible<2->{
\item {\color{blue} Centralized}: $M^n = (M_1,\ldots,M_n) = f(X_1,\ldots,X_n)$ }
\visible<3->{
\item {\color{red} Adaptive / Sequential}: $M_i = f_i(X_i,M^{i-1})$ }
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Related Work}
\begin{itemize} 
\pause
\item Estimation via compressed information [Han '87], [Zhang \& Berger '88] (centralized)
\pause
\item Estimation from multiple machines subject to a bit constraint [Zhang, Duchi, Jordan, Wainwright '13] (distributed / adaptive)
\pause
\item Distributed hypothesis testing under quantization [Tsitsiklis '88] (distributed)
\pause
\item Remote multiterminal source coding (CEO) [Berger, Zhang, Wiswanathan '96], [Oohama '97] (distributed)
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Consistency}
Q: in what setting consistent estimation is possible? \\
\bigskip
\pause
A: all ! 
\pause
\bigskip
\textbf{Proof:}
\[
M_i = \mathbf 1(X_i>0),\quad i=1,\ldots,n
\]
(distributed setting)
\pause
\[
\frac{1}{n}\sum_{i=1}^n M_i \rightarrow \Prob\left(X>0 \right) = \Phi\left(\thetac / \sigma \right)
\]
\end{frame}

%SAY: supper slow, of course. Can we do better? what is the best possible?

@@

\begin{frame}
\frametitle{Efficiency}
\textbf{Definition:} \emph{asymptotic relative efficiency (ARE)}:
\[
\mathrm{ARE}(\widehat{\theta}_n) \triangleq \lim_{n\rightarrow \infty} \mathbb E \left[ \left(\widehat{\theta}_n-\theta \right)^2 \right] / \left(\sigma^2 / n \right)
\]
(relative to minimax risk without bit constraint)

%SAY: many definitions for ARE are possible, we use this one becasue our results apply to all of them

@@

\begin{prop} ARE under centralized encoding is $1$
\end{prop}
\pause

\textbf{Proof:}
\[
\mathbb E \left( \theta- \widehat{\theta} \right)^2 = \overbrace{\mathbb E \left( \theta- \bar{\theta} \right)^2}^{\sigma^2/n} + \underbrace{\mathbb E \left( \bar{\theta}- \widehat{\theta} \right)^2}_{O(2^{-2n})}
\]
\pause
\begin{itemize}
\item Encoder is required to describe $\bar{\theta}$ using $n$ bits
\begin{itemize}
\item divide parameter space $\Theta$ into $2^n$ regions of equal size
\item send region index where $\bar{\theta}$ falls
\end{itemize}
\end{itemize}
\bigskip
\pause
Note: mean of $\bar{\theta}$ is unknown -- hard to derive a globally optimal strategy
%SAY: exploration -- exploitation even in this simple setting
\end{frame}

@@

\begin{center}
\vspace{-20pt}
\begin{tikzpicture}[node distance=2cm,auto,>=latex]
  \node at (0,0) (source) {$\thetac_j$} ;
  \node[ssum, right of = source, scale = 0.75] (sum1) {+};
  \node[above of = sum1, node distance = 1cm] (z1) {$Z_{1}$};
  \node[int1, right of = sum1, node distance = 2cm, scale = 0.75] (enc1) {Enc 1};  
\draw[->,line width = 2pt] (source) -- (sum1); 
\draw[->,line width = 2pt] (sum1) -- node[above] {$X_{1}$} (enc1); 
\draw[->,line width = 2pt] (z1) -- (sum1);

@@

 \node[coordinate,below of = source, node distance = 2.5cm] (source2) {};
 \node[ssum, right of = source2, scale = 0.75] (sum2) {+};
  \node[above of = sum2, node distance = 1.2cm] (z2) {$Z_{n}$};
\node[int1, right of = sum2, node distance = 2cm, scale = 0.75] (enc2) {Enc n};

@@

\draw[->,line width = 2pt] (sum2) -- node[above] {$X_{n}$} (enc2); 
\draw[->,line width = 2pt] (z2) -- (sum2);

@@

%\node[int1, right of = enc2, node distance = 6cm, scale = 0.75] (est) {Est};
\draw[->,line width = 2pt] (enc1) -- node[above, xshift = 0cm, rotate = -15] (mes1) {one bit} (est);   
\draw[->,line width = 2pt] (enc2) -- node[above, xshift = 0cm, rotate = 12] (mes2) {one bit} (est);   
\node[right of = est, node distance = 1.5cm] (dest) {$\widehat{\thetac}_j$};
\draw[->, line width=2pt] (est) -- (dest);
\end{tikzpicture}
\end{center}
\pause
Assume:
\begin{itemize}
\item $Z_{1},\ldots,Z_{n} \overset{\mathrm{i.i.d}}{\sim}\mathcal N(0,\sigma^2)$
\pause
\item Replicate $k$ times: $\thetac_1,\ldots,\thetac_k \overset{\iid}{\sim} \mathcal N(0,\sigma_0^2)$ 
\pause
\item Block encode $X_{j,1},\ldots,X_{j,n}$ using $k$ bits, $j=1,\ldots,k$
\pause
\item $D_{CEO} = \inf_{k,\widehat{\theta}} \frac{1}{k} \sum_{j=1}^k \mathbb E \left( \theta_j - \widehat{\theta}_j\right)^2$
\end{itemize}
%SAY: lower bound to our setting
\end{frame}

\begin{frame}
Quadratic Gaussian CEO under optimal rate allocation [Chen, Zhang, Berger, Wicker '04] :
\[
 D_{CEO}  \geq \frac{4}{3} \frac{\sigma^2}{n} + o(1/n) 
\]
%\begin{itemize}
%\item[] CEO distortion satisfies [Chen, Zhang, Berger, Wicker '04]:
%\[ D^\star = \frac{4}{3} %\frac{\sigma^2}{n} + o(n^{-1}) \]
%\end{itemize}
\pause
\begin{alertblock}{Conclusion}
Distributed encoding will hurt you (even if you can repeat experiment and encode over blocks)
\end{alertblock}
\bigskip
\pause
\begin{itemize}
\item In fact [K., Rini, Goldsmith '17]:
\end{itemize} 
\[
 D_{CEO} \leq \frac{4}{3} \frac{\sigma^2}{n} + \frac{\sigma_0^2}{3n} + o(1/n) 
\]
(ARE of $4/3$ is achievable with coding over blocks)
%SAY: depending on how large our bound from $k=1$, it tells us that distributed setting is not the limitng factor
\end{frame}

@@

\node[below of = source2, node distance = 2cm] (source3) {$X_n$};
\node[int1, right of = source3, node distance = 2cm, scale = 0.75] (enc3) {Enc n};  
\draw[->,line width = 2pt] (source3) -- (enc3);

%\node[above of = source3, node distance = 1cm, xshift = -0.5cm] (source3) {$X_i \sim \mathcal N(\theta,\sigma^2)$};

\node[above of = source, node distance = 0.75cm] (dist) {$X_i \sim {\mathcal N} \left(\theta, \sigma^2 \right)$};

\node[Est, right of = enc2, node distance = 4.5cm] (est) {Est};
\draw[->,line width = 2pt] (enc1) -| node[above, xshift = -2.5cm] (mes1) {$M_1 \in \left\{-1,1\right\}$} (est);   
\draw[->,line width = 2pt] (enc2) -- node[above, xshift = 0cm] (mes2) {$M_2 \in \left\{-1,1\right\}$} (est);   
\draw[->,line width = 2pt] (enc3) -| node[above, xshift = -2.5cm]  {$M_n \in \left\{-1,1\right\}$} (est);   
\node[right of = est, node distance = 2cm] (dest) {$\widehat{\theta}(M^n)$};
\draw[->, line width=2pt] (est) -- (dest);

\visible<2>{
\draw[<->,line width = 2pt] (enc1) -- (enc2);   
\draw[<->,line width = 2pt] (enc2) -- (enc3);
\draw[dashed,line width = 2pt, color = white] (enc2)+(0,-0.5) -- +(0,-1.5);}

\visible<3->{
\draw[->,line width = 2pt, color = red] (enc1)+(1,0) -- (enc2);   
\draw[->,line width = 2pt, color = red] (enc2)+(1,0) -- (enc3);
\draw[line width = 2pt, color = red] (enc1)+(1,0) -- +(1,-1.5);

\draw[dashed,line width = 2pt, color = white] (enc2)+(0.66,-0.7) -- +(0.25,-1.5);}

\begin{itemize}
\item Distributed: $M_i = f_i(X_i)$  \hfill{ARE $\geq 4/3$}
\visible<2->{
\item {Centralized}: $M^n = f(X_1,\ldots,X_n)$ \hfill{ARE $=1$} }
\visible<3->{
\item {\color{red} Adaptive / Sequential: $M_i = f_i(X_i,M^{i-1})$ \hfill{ARE$=\pi/2$} }  }
\end{itemize}
\end{frame}

@@

\begin{theorem}[achievability]
There exists an estimator with ARE $\pi/2$
\end{theorem}

@@

%\begin{theorem}[one-step optimal strategy]
%The next step one-bit message that minimizes the MSE is of the form $M = \sgn (X-\tau)$ where $\tau$ satisfies the fixed-point equation 
%\[
%\tau = \frac{1}{2} \left( \frac{ \int_{-\infty}^\tau \theta \pi(d\theta) }{\int_{-\infty}^\tau \pi(d\theta)}  + \frac{ \int_{\tau}^\infty \theta \pi(d\theta)}{\int_{\tau}^\infty \pi(d\theta)} \right)
%\]
%\end{theorem}
\end{frame}

\begin{frame}
\frametitle{Achievability}
\framesubtitle{existence of an estimator with ARE $= \pi/2$}

\begin{itemize}
\item[(i)] For $X \sim \mathcal{N}(\theta,\sigma^2)$, $\mathsf{med}(X) = \theta$ 
\pause
\item[(ii)] $\mathsf{med}(X) = \underset{m}{ \mathrm{argmin~}} \mathbb E \left|X-m \right|$
\pause
\item[(iii)] Stochastic gradient descent on $\mathbb E \left| X-\theta \right|$:
\[
\theta_n = \theta_{n-1} + \gamma_n \sgn(X_n - \theta_n)
\]
\[
\widehat{\theta}_n = \frac{1}{n} \sum_{i=1}^n \theta_i 
\]
\end{itemize}
\bigskip
\pause
From [Polyak \& Juditsky '92] (under conditions on $(\gamma_n)$): 
\[
\sqrt{n}(\theta- \widehat{\theta}_n) \rightarrow \mathcal N\left(0, \sigma^2 \pi /2 \right)
\]
\end{frame}
%picture

%\begin{frame}
%\frametitle{Achievability} 
%
%[Polyak \& Juditsky '92]: \\
%
%
%where:
%\begin{itemize}
%\item[(i)] $\gamma_n \rightarrow 0^+$ ``not too slow'' 
%\item[(ii)] $\psi(x) = \mathbb E \varphi(x+Z)$ 
%\item [(iii)] $\chi(x) = \mathbb E \varphi^2(x+Z)$ 
%\item[(iv)] some regularity  conditions on $\varphi$, $\chi$, $\psi$
%\end{itemize}
%Then 
%\[
%\sqrt{n}(\theta- \widehat{\theta}) \rightarrow \mathcal N\left(0,V\right)
%\]
%where $V = \chi(0)/\psi'^2(0)$.  \\
%\medskip
%\pause
%Proof of theorem: take $\varphi(x) = \sgn(x)$
%\end{frame}

\begin{frame}
\frametitle{Converse}
\framesubtitle{ARE $\geq \pi/2$}
The van-Trees inequality (e.g. [Tsybakov '08]) implies
\[
\mathbb E \left(\theta - \widehat{\theta} \right)^2 \geq \frac{1}{\mathbb E I_\theta(M^n) + I_\pi} \geq \frac{1}{\sum_{i=1}^n I_\theta(M_i|M^{i-1}) + I_\pi} 
\]
$I_\pi$ is the location Fisher information with w.r.t. some prior $\pi(d\theta)$ on $\Theta$\\

\begin{lem}[K. \& Duchi '17]
\[
I_\theta(M_i|M^{i-1}) \leq \frac{2}{\pi\sigma^2}
\]
\end{lem}
\textbf{Proof}: \\
Stein identity implies that portion maximizing the information is a threshold: $M_i^{-1}(1) = (\theta,\infty)$  \\
The rest follows by induction over number of portions
%SAY: the rest follows by induction over number 
\end{frame}

%\begin{frame}
%\frametitle{One-step Optimal Bit}
%% SAY: previously we bound the inforamtion under any scheme. In addition, it is interesting to understand what is the one-step optimal scheme
%%Q: upon observing $M^{n-1}$, what is the optimal detection region $M_n^{-1}(1)$ that minimizes the MSE ? \\ \pause
%\begin{theorem}
%Let $\pi(\theta)$ be an absolutely continuous log-concave distribution. For $X\sim \mathcal N(\theta,\sigma^2)$ let
%\[
%M = \sgn(X-\tau)
%\]
%where $\tau$ is the unique solution to
%\[
%\tau = \frac{1}{2} \left( \frac{ \int_{-\infty}^\tau \theta \pi(d\theta) }{\int_{-\infty}^\tau \pi(d\theta)}  + \frac{ \int_{\tau}^\infty \theta \pi(d\theta)}{\int_{\tau}^\infty \pi(d\theta)} \right)
%\]
%Then for any $M'(X) \in \{-1,1\}$ and $\widehat{\theta}(M')$:
%\[
%\mathbb E \left(\theta- \widehat{\theta}(M') \right)^2 \geq \mathbb E \left(\theta - \mathbb E[\theta|M] \right)^2
%\]
%\end{theorem}
%\vspace{-20pt}
%\pause
%\begin{alertblock}{Interpertation:}
%\begin{itemize}
%\item One-bit message that minimizes MSE is a threshold detector
%\pause
%\item Leads to a Bayes procedure $\mathbb E \left[\theta | M\right]$
%\end{itemize}
%\end{alertblock}
%
%%SAY: we don't know if this procedure is optimal, what we conjecture that it is
%\end{frame}

%\begin{frame}
%\frametitle{One-step Optimal Scheme}
%Initialization: $P_0(t) = \pi(\theta)$ \\
%Repeat for $n \geq 1$:
%\begin{itemize}
%\item[(i)] $P_n(t) = \Prob(\theta = t|M^n) = \alpha_n P_{n-1}(t) \Phi \left(M_n \frac{t-\tau_{n-1}}{\sigma} \right)$ 
%\item[(ii)] $\widehat{\theta} = \mathbb E[\theta | M^n] = \int t P_n(t)dt $ 
%\item[(iii)] Find $\tau_n$ from 
%\[
%\tau_n = \frac{1}{2} \left( \frac{ \int_{-\infty}^\tau t P_n(t)dt }{\int_{-\infty}^\tau P_n(t)dt}  + \frac{ \int_{\tau}^\infty t P_n(t)dt}{\int_{\tau}^\infty P_n(t)dt} \right)
%\]
%\item[(iv)] $M_{n+1} = \sgn(X_{n+1}-\tau_n)$
%\end{itemize}
%\end{frame}

%\begin{frame}
%\frametitle{Numerical Example}
%\begin{center}
%Normalized empirical risk versus number of samples $n$ \\
%(1000 Monte Carlo experiments)
%\begin{tikzpicture}
%\node at (0,0) {\includegraphics[scale=0.5]{one_bit_adaptive}};
%%\node[color = blue,  minimum width=1cm, align = left] at (1.5,1.5) {assymptotically optimal \eqref{eq:sgd_est}};
%%\node[color = red, minimum width=1cm, align = left] at (1.5,0.55) {one-step optimal \eqref{eq:message_update}} ;
%\node[rotate = 90, scale = 1] at (-5,0) {$n \mathbb E \left(\widehat{\theta}_n - \theta \right)^2$};
%\node[scale = 1] at (0,-3) {$n$};
%%\draw[dashed] (-3.5,1.18) node[left, scale = 0.7] {$\frac{\pi}{2}$}-- (-3,1.18);
%\end{tikzpicture}
%$\theta \sim \unif(-3,3)$
%\end{center}
%\end{frame}

@@
